---
title: "Master Notebook for Digitization"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: github_document
---

## Introduction
This notebook executes all of the participant-level digitization notebooks. Maps with all of the raw data can be found on the project's [Open Science Framework repository](<REDACTED FOR PEER REIVEW>).

## Dependencies
This notebook requires the following packages:

```{r load-packages}
# tidyverse packages
library(purrr)
library(rmarkdown)

# other packages
library(here)
```

## Create Vector of Files
All of the files contained in `docs/` can be listed and saved in a vector:

```{r identify-files}
# create vector of all files
files <- dir(path = here("docs"), pattern = "*.Rmd")

# isolate participant files
participant_files <- grep("^participant", files, value = TRUE)
```

## Knit Participant Documents
With the list of participant files created, we can iterate over them and knit each one:

```{r render-participants}
# render
participant_files %>%
  map(~render(input = here("docs", .x), output_format = "github_document"))
```

The `.html` output is not needed, and so we'll remove it from the `docs/` folder:

```{bash clean-docs}
rm *.html
```

And clean-up our global enviornment:

```{r clean-up-enviornment}
rm(files, participant_files)
```

