---
title: "Analyze Clusters"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: github_document
---

## Introduction
This notebook calculates Moran's $I$ statistics for both the frequent and regular data.

## Dependencies
This notebook requires the following packages:

```{r load-packages}
# tidyverse packages
library(dplyr)

# spatial packages
library(spdep)
library(sf)

# other packages
library(here)
```

## Load and Prepare Data
We'll load our data and papre it by converting it to an `sp` object. First, the frequent data:

```{r prep-frequent}
freq_respondents <- st_read(here("data", "summaries", "freq_respondents"), stringsAsFactors = FALSE) %>%
  st_transform(crs = 26915)

freq_respondents <- mutate(freq_respondents, frequent_pct = frequent/53*100)

freq_respondents <- as(freq_respondents, Class = "Spatial")
```

Then the regular data:

```{r prep-regular}
reg_respondents <- st_read(here("data", "summaries", "reg_respondents"), stringsAsFactors = FALSE) %>%
  st_transform(crs = 26915)

reg_respondents <- mutate(reg_respondents, regular_pct = regular/53*100)

reg_respondents <- as(reg_respondents, Class = "Spatial")
```

## Analyze Frequent Data
We need to create a spatial weights object (using rook) for our data:

```{r freq-weights}
nb <- poly2nb(freq_respondents, queen=FALSE)
lw <- nb2listw(nb, style="W", zero.policy=TRUE)
```

With a weights object created, we can run a Moran's I test:

```{r freq-moran}
moran.test(freq_respondents$frequent_pct,lw)
```

And then finally use a monte carlo simulation to estimate the $p$ value for our test:

```{r freq-sim}
moran.mc(freq_respondents$frequent_pct, lw, nsim=999)
```

## Analyze Regular Data
We need to create a spatial weights object (using rook) for our data:

```{r reg-weights}
nb <- poly2nb(reg_respondents, queen=FALSE)
lw <- nb2listw(nb, style="W", zero.policy=TRUE)
```

With a weights object created, we can run a Moran's I test:

```{r reg-moran}
moran.test(reg_respondents$regular_pct,lw)
```

And then finally use a monte carlo simulation to estimate the $p$ value for our test:

```{r reg-sim}
moran.mc(reg_respondents$regular_pct, lw, nsim=999)
```

## Clean-up Enviornment
Finally, we'll clean-up our enviornment:

```{r clean-up}
rm(freq_respondents, reg_respondents, lw, nb)
```

