---
title: "Participant 149"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: github_document
params:
  pid: 149
---

## Introduction
This notebook is used for digitizing participant data for the `qualmap` pilot project.

## Dependencies
This notebook requires the following packages:

```{r load-packages}
# spatial packages
library(qualmap)
library(sf)

# tidyverse packages
library(readr)

# other packages
library(here)
library(testthat)
```

## Load Reference Data
First, we'll load the reference data from `data/grids/`:

```{r load-data}
grids <- st_read(here("data", "grid"), stringsAsFactors = FALSE)
```

## Define and Validate Participant Clusters
First, we'll define the clusters associated with the "frequent" prompt:

```{r define-frequent-clusters}
cluster1 <- qm_define("C-2", "D-2")
cluster2 <- qm_define("F-3", "G-3", "G-4")
cluster3 <- qm_define("G-3")
cluster4 <- qm_define("F-5", "F-6", "E-6")
```

Next, we'll validate these data. The validation is performed in the context of a unit test to ensure that each object is valid:

```{r validate-frequent-clusters}
expect_equal(qm_validate(ref = grids, key = GRID_ID, value = cluster1), TRUE)
expect_equal(qm_validate(ref = grids, key = GRID_ID, value = cluster2), TRUE)
expect_equal(qm_validate(ref = grids, key = GRID_ID, value = cluster3), TRUE)
expect_equal(qm_validate(ref = grids, key = GRID_ID, value = cluster4), TRUE)
```

No output on these unit tests means that each cluster is indeed valid.

Next, we'll repeat the process with the "regular" prompt by first entering the cluster data:

```{r define-regular-clusters}
cluster5 <- qm_define("D-2", "E-2")
```

As before, we'll validate these data in the context of a unit test where no output means each is valid:

```{r validate-regular-clusters}
expect_equal(qm_validate(ref = grids, key = GRID_ID, value = cluster5), TRUE)
```

## Create Cluster Objects
With all data validated, we can now convert each to cluster object. This process adds relevant metadata about each cluster:

```{r create-cluster-objects, warning=FALSE}
cluster1 <- qm_create(ref = grids, key = GRID_ID, value = cluster1, rid = params$pid, cid = 1, category = "frequent")
cluster2 <- qm_create(ref = grids, key = GRID_ID, value = cluster2, rid = params$pid, cid = 2, category = "frequent")
cluster3 <- qm_create(ref = grids, key = GRID_ID, value = cluster3, rid = params$pid, cid = 3, category = "frequent")
cluster4 <- qm_create(ref = grids, key = GRID_ID, value = cluster4, rid = params$pid, cid = 4, category = "frequent")
cluster5 <- qm_create(ref = grids, key = GRID_ID, value = cluster5, rid = params$pid, cid = 5, category = "regular")
```

## Combine Cluster Objects
With all of the cluster objects created, we can now combine them into a single object:

```{r combine}
clusters <- qm_combine(cluster1, cluster2, cluster3, cluster4, cluster5)
```

## Write Combined Object
Finally, we'll write these data to a `.csv` file:

```{r write-csv}
write_csv(clusters, here("data", "participants", paste0("participant_", params$pid, ".csv")))
```
