---
title: "Summarize Clusters"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output: github_document
---

## Introduction
This notebook imports all of the raw cluster data and summarizes it into mappable `sf` objects.

## Dependencies
This notebook requires the following packages:

```{r load-packages}
# spatial packages
library(qualmap)
library(sf)

# tidyverse packages
library(dplyr)
library(purrr)
library(readr)
library(tidyr)

# other packages
library(here)
```

## Load Data
First, we need to create a list of the `.csv` files to be read in:

```{r create-file-list}
files <- dir(path = here("data", "participants"), pattern = ".csv")
```

Next, we can read them in and reduce them to a single data frame with `purrr`:

```{r load-data}
data <- files %>%
  map_df(~read_csv(file = here("data", "participants", .x)))
```

Finally, we'll load the reference data from `data/grids/`:

```{r load-ref-data}
grids <- st_read(here("data", "grid"), stringsAsFactors = FALSE)
```

## Modify Grid
Not all grid squares were shown on the instrument, so we'll remove some of the outer grid squares:

```{r}
grids %>% 
  separate(GRID_ID, into = c("col", "row"), sep = "-", remove = FALSE) %>%
  filter(col != "A") %>%
  filter(col != "I") -> grids
```


## Verify
Next, we can use the `qualmap` function `qm_verify()` to ensure we have data that are appropriate for summarization:

```{r verify}
data <- qm_verify(data)
```

`qm_verify` will error if there is an issue. Otherwise, it returns a tibble with the class `"qm_cluster"` applied to it, which is necessary for using `qm_summarize()` below.

## Summarize
The `qualmap` package comes with two options for counting, by cluster and respondent. Theoretically, a feature could be included in multiple clusters by a single respondent. Setting `count = "clusters"` will return a count of the number of clusters that a feature is included in, whereas `count = "respondents"` returns a count of the number of respondents who identified that feature in at least one cluster.

We'll create output using both options for the "frequent" qualitative construct:

```{r frequent}
freq_clusters <- qm_summarize(ref = grids, key = "GRID_ID", clusters = data, category = "frequent", count = "clusters")
freq_respondents <- qm_summarize(ref = grids, key = "GRID_ID", clusters = data, category = "frequent", count = "respondents")
```

We'll also create output using both options for the "regular" qualitative construct:

```{r regular}
reg_clusters <- qm_summarize(ref = grids, key = "GRID_ID", clusters = data, category = "regular", count = "clusters")
reg_respondents <- qm_summarize(ref = grids, key = "GRID_ID", clusters = data, category = "regular", count = "respondents")
```

## Write Data
Finally, we'll write each of the files to `data/` as shapefiles:

```{r write-data}
st_write(freq_clusters, here("data", "summaries", "freq_clusters", "freq_clusters.shp"), delete_dsn = TRUE)
st_write(freq_respondents, here("data", "summaries", "freq_respondents", "freq_respondents.shp"), delete_dsn = TRUE)
st_write(reg_clusters, here("data", "summaries", "reg_clusters", "reg_clusters.shp"), delete_dsn = TRUE)
st_write(reg_respondents, here("data", "summaries", "reg_respondents", "reg_respondents.shp"), delete_dsn = TRUE)
```
